Fractional Decomposition Tree Algorithm: A tool for studying the integrality gap of Integer Programs

### 2020-08-08 FDT2020/main-section.tex
arash comments in page 20

As described in Section~\ref{sec:2EC-intro}), when the LP optimum for (\ref{eq:subtour}) is a Carr-Vempala point (see Figure~\ref{fig:CVpoint}), the integrality gap of that instance is equal to that of 2EC. So in that sense, these are the hardest fractional points to round to a feasible solution. We now compare the quality of FDT's solution from a Carr-Vempala point to that of Christofides, the best known approximation algorithm.  In the classic graph-based Christofides algorithm for TSP, find a minimum spanning tree of the graph.  Then, since TSP runs on a complete graph, add a perfect matching on the vertices that have odd degree in the spanning tree to make the (multi)graph 2-connected. If we stop that algorithm without shortcutting, we have a feasible solution for TSP. For 2EC, however, the graph is not usually complete. We can instead add an $O$-join, where $O$ is the set of odd-degree vertices in the spanning tree~\footnote{For graph $G=(V,E)$ and $O\subseteq V$ with $|O|$ even, an $O$-join of $G$ is a subgraph of $G$ that has odd degree on the vertices in $O$ and even degree on vertices in $V\setminus O$. Graphs always have an even number of odd vertices since every edge has two endpoints, giving an even number of endpoints.}. The best-of-many Christofides (BOMC) algorithm~\cite{AKS15} works better than the classic Christofides algorithm in practice~\cite{GW17}.  The BOMC algorithm samples many spanning trees from the LP relaxation (with edges chosen proportionally to the LP value), augments each spanning tree with the matching, and taking the best overall solution.

We now describe the polyhedral version of Christofides' algorithm~\cite{Wolsey} specifically for Carr-Vempala points.
% Consider the LP~(\ref{eq:subtour}) in~(\ref{eq:subtour}) and any feasible point $x$. 
%Consider any partitioning of the vertices of the graph into $b$ pieces and let $E_c$ be the set of edges that connect any two of the $b$ pieces. Then we have $\sum_{E_c} x_e \ge b$. That is because there are (fractionally) at least $2$ edge endpoints in each piece, for a total of at least $2b$ endpoints, and at least $b$ edges. An LP relaxation for the spanning-tree polytope will guarantee $b-1$ edges connecting any $b$ groups of vertices.  Thus we can scale any LP solution for the subtour-elimination polytope for an $n$-node graph by $(n-1)/n$ and it will still be feasible for the spanning tree polytope. 
%\cindy{I think it would be very helpful for the reader for you to remind them of the classic (graph-based, not polyhedral) metric (triangle-inequality) Christofides algorithm here and argue that it applies to 2EC.  Most people will remember Christofides' TSP approximation algorithm because it's so classic.  But it's easy to remind people.  Then it would be helpful to explain that when you give the LP in (\ref{ojoinaverage}) how that relates to Christofides.  If this paper is accepted, all kinds of CS/OR researchers may want to read it, so we should make their job as easy as possible.  It would be nice for others to use FDT and to build off it.}
%It is well known that $\frac{|V|-1}{|V|}\cdot x$ is in the convex hull of incidence vectors of spanning trees of $G$.
%\cindy{Is this for any feasible solution, or extreme point of the subtour LP or just Carr-Vempala points?  It seems that it's for any feasible point since you don't remind the reader about Carr-Vempala points until later.}
%\cindy{The statement before this depends very much on the audience.  For those who study polyhedral combinatorics especially for TSP and who know what a Carr-Vempala point is off the top of their heads, it may be well known.  For people who generally do combinatorial optimization, I would not expect them to know this. So this kind of phrasing might annoy a reader.  General rule: either a statement is really obvious (so just state it), or it is not (so give a reference or justification). Otherwise (and I know you don't intend this), you are implying the reader is stupid if they don't see the argument right away. Of course, it's the authors' job to teach and convince the reader. Thus one should avoid phrases like ``It's obvious/trivial that...'' and similar phrases. I would change the sentence before this comment to ``The vector $\frac{|V|-1}{|V|}\cdot x$ is in the convex hull...'' and give a short justification or reference. Or ``Foo and Barr (cite Foo and Bar's paper) argue that ... is in the convex hull...'' I'm sure I could figure this out if a had a little time to think, but I don't have that time now.}
%\cindy{I'm going to remove things like ``It's easy to see that,'' where I can. But we may need to add a justification or reference.}
Let $x$ be a Carr-Vempala point (from~(\ref{eq:subtour})) defined on a graph $G=(V,E)$.  From~\cite{Vygen12}, we can decompose a slightly-scaled-down $x$ into a convex combination of spanning trees:
\begin{equation}
\label{eq:tree-decomp}
\frac{|V|-1}{|V|}\cdot x= \sum_{i=1}^{k}\lambda_i\chi^{T_i},
\end{equation}
where $T_i$ is spanning tree of $G$, $\chi^{T_i}$ is its incidence vector, $\sum_{i=1}^{k}\lambda_i=1$, and $\lambda_i\geq 0$ for $i\in [k]$. The spanning-tree polytope is integral so there are polynomial-time algorithms to construct this decomposition (See Proposition~\ref{cara} and surrounding discussion).

Let $O_i$ be the set of odd-degree vertices of spanning tree $T_i$.
%Consider the incidence vectors of feasible $O_i$-joins.
% \cindy{Arash, do you also want to replace this longer discussion with a reference?}
Adding an $O_i$-join to tree $T_i$ gives a solution to 2EC.
% CAP: TODO, use later?
%Let $S$ be any subset of vertices with an even number of odd-degree vertices.  The $O_i$-join LP requires at least one edge joining $S$ to the rest of the vertices. The subtour-elimination LP requires every subset of vertices have two edges connecting it to the rest of the vertices.  Therefore, since $x$ is a feasible solution to the subtour-elimination LP, $x/2$ is feasible for the $O_i$-join LP associated with spanning tree $T_i$.
We will use Wolsey's observation~\cite{wolsey} that $\frac{x}{2}$ is in the convex hull of incidence vectors of the $O_i$-join of $G$ (denoted $O_i\join(G)$).
%\cindy{There was another instance of ``It's easy to see that...'' Given that (quite reasonably) you do not expect the reader to know what an O-join is, you should probably just give the short justification for this statement and remove ``It's easy to see that.''}
%\cindy{Do you have to precompute the joins (If they are more than the spanning trees themselves)?  Does that mean generating spanning trees before setting up the LP?  Or generating them on the fly?  Only minimum-weight spanning trees?  It seems that with a little more explanation, a reader who has not been immersed in 2EC and TSP could follow this.}
We solve the following LP that allows us to find edges that simultaneously augment all the spanning trees to solutions to 2EC.
%$O$-join-based parity corrections that are good for the whole convex combination.
\begin{equation}\label{ojoinaverage}
\min \{ \alpha:\;\sum_{i=1}^{k} \lambda_i y^i = \alpha \cdot x,\;  y^i \in O_i\join(G) \; \mbox{for $i\in [k]$}\}.
\end{equation}
The variables in the above LP are $y^i\in \mathbb{R}^{E}_{\geq 0}$ for $i\in [k]$, where $E$ is the number of edges in the graph.  The parameters $\lambda_i$ are those we computed in~(\ref{eq:tree-decomp}). For each $i\in [k]$, we require $y^i$ to be in the dominant of $O_i\join(G_x)$.  We enforce this by adding a set of constraints for each $i$ from~\cite[p. 490]{schrijverbook}. These constraints force the inclusion of at least one edge from the cut associated with each set of vertices that have an even number of elements from $O_i$. The $O_i$ joins are selected fractionally, but because the $O_i$-join polytope is integral, we can find integral solutions (proposition~\ref{cara}).  This method gives a ($\frac{|V|-1}{|V|}+\alpha$)-approximation for the specific instance $x$. In general this can be less than the classic $\frac{3}{2}$.
%\cindy{You say the $y^i$ are LP variables, not integers. It even looks like they can even be more than $1$. What does it mean for a non-binary vector to be in a set of binary variables?  I would have expected the lambdas to be the variables, given their use as the convex combination multipliers in most of the paper.}
%This formulation allows the instance-specific approximation ratio of Christofides' algorithm to be below $\frac{3}{2}$.
\cindy{Arash, I liked your explanation about how a single edge will need to contribute 3/2 in the classic method.  You can add that if you'd like, but that might not be necessary given the time.}

>>>>

As described in Section~\ref{sec:2EC-intro}, when the LP optimum for (\ref{eq:subtour}) is a Carr-Vempala point (see Figure~\ref{fig:CVpoint}), the integrality gap of that instance is equal to that of 2EC. So in that sense, these are the hardest fractional points to round to a feasible solution. We now compare the quality of FDT's solution from a Carr-Vempala point to that of Christofides, the best known approximation algorithm.  In the classic graph-based Christofides algorithm for TSP, find a minimum spanning tree of the graph.  Then, since TSP runs on a complete graph, add a perfect matching on the vertices that have odd degree in the spanning tree to make the (multi)graph 2-edge-connected. If we stop that algorithm without shortcutting, we have a feasible solution for TSP. \arash{The following sentence is not accurate, Metric-TSP is the same as TSP on a non-complete graph where we can take multiple copies of the edges, Metric-2EC (given complete graph with metric costs find the minimum cost 2-edge-connected SUBgraph -no doubling of edges-) is the same as 2EC on a non-complete graph where we can take multiple copies of edges.}
For 2EC, however, the graph is not usually complete. We can instead add an $O$-join, where $O$ is the set of odd-degree vertices in the spanning tree~\footnote{For graph $G=(V,E)$ and $O\subseteq V$ with $|O|$ even, an $O$-join of $G$ is a subgraph of $G$ that has odd degree on the vertices in $O$ and even degree on vertices in $V\setminus O$. Graphs always have an even number of odd vertices since every edge has two endpoints, giving an even number of endpoints.}. The best-of-many Christofides (BOMC) algorithm~\cite{AKS15} works better than the classic Christofides algorithm in practice~\cite{GW17}.  The BOMC algorithm samples many spanning trees from the LP relaxation (with edges chosen proportionally to the LP value), augments each spanning tree with the matching, and taking the best overall solution.

\arash{I think here we should say that something like this: the polyhedral analaysis of Christofides shows that if $x\in \subtour(G)$, then $\frac{3}{2}x$ dominates a convex combination of connected Eulerian (hence 2-edge-connected) multigraph of $G$ \cite{wolsey,shmoyswilliamson}. This bound holds for any point $x\in \subtour(G)$ for any graph $G$, including any Carr-Vempala point. Below we ask if this bound can be improved for specific instances by doing a slightly improved analysis of the algorithm}\arash{Then we explain the LP in (32)}

We now describe the polyhedral version of Christofides' algorithm~\cite{Wolsey} specifically for Carr-Vempala points.
% Consider the LP~(\ref{eq:subtour}) in~(\ref{eq:subtour}) and any feasible point $x$. 
%Consider any partitioning of the vertices of the graph into $b$ pieces and let $E_c$ be the set of edges that connect any two of the $b$ pieces. Then we have $\sum_{E_c} x_e \ge b$. That is because there are (fractionally) at least $2$ edge endpoints in each piece, for a total of at least $2b$ endpoints, and at least $b$ edges. An LP relaxation for the spanning-tree polytope will guarantee $b-1$ edges connecting any $b$ groups of vertices.  Thus we can scale any LP solution for the subtour-elimination polytope for an $n$-node graph by $(n-1)/n$ and it will still be feasible for the spanning tree polytope. 
%\cindy{I think it would be very helpful for the reader for you to remind them of the classic (graph-based, not polyhedral) metric (triangle-inequality) Christofides algorithm here and argue that it applies to 2EC.  Most people will remember Christofides' TSP approximation algorithm because it's so classic.  But it's easy to remind people.  Then it would be helpful to explain that when you give the LP in (\ref{ojoinaverage}) how that relates to Christofides.  If this paper is accepted, all kinds of CS/OR researchers may want to read it, so we should make their job as easy as possible.  It would be nice for others to use FDT and to build off it.}
%It is well known that $\frac{|V|-1}{|V|}\cdot x$ is in the convex hull of incidence vectors of spanning trees of $G$.
%\cindy{Is this for any feasible solution, or extreme point of the subtour LP or just Carr-Vempala points?  It seems that it's for any feasible point since you don't remind the reader about Carr-Vempala points until later.}
%\cindy{The statement before this depends very much on the audience.  For those who study polyhedral combinatorics especially for TSP and who know what a Carr-Vempala point is off the top of their heads, it may be well known.  For people who generally do combinatorial optimization, I would not expect them to know this. So this kind of phrasing might annoy a reader.  General rule: either a statement is really obvious (so just state it), or it is not (so give a reference or justification). Otherwise (and I know you don't intend this), you are implying the reader is stupid if they don't see the argument right away. Of course, it's the authors' job to teach and convince the reader. Thus one should avoid phrases like ``It's obvious/trivial that...'' and similar phrases. I would change the sentence before this comment to ``The vector $\frac{|V|-1}{|V|}\cdot x$ is in the convex hull...'' and give a short justification or reference. Or ``Foo and Barr (cite Foo and Bar's paper) argue that ... is in the convex hull...'' I'm sure I could figure this out if a had a little time to think, but I don't have that time now.}
%\cindy{I'm going to remove things like ``It's easy to see that,'' where I can. But we may need to add a justification or reference.}
Let $x$ be a Carr-Vempala point (from~(\ref{eq:subtour})) defined on a graph $G=(V,E)$. \arash{I think it's better to say that: It is a well-known fact that if $x\in \subtour(G)$, then $\frac{|V|-1}{|V|}x$ can be written as a convex combination of spanning trees of $G$ (See Proposition 2 in \cite{Vygen12} for instance), the reason is that this is because Vygen didn't come up with this.} From~\cite{Vygen12}, we can decompose a slightly-scaled-down $x$ into a convex combination of spanning trees:
\begin{equation}
\label{eq:tree-decomp}
\frac{|V|-1}{|V|}\cdot x= \sum_{i=1}^{k}\lambda_i\chi^{T_i},
\end{equation}
where $T_i$ is spanning tree of $G$, $\chi^{T_i}$ is its incidence vector, $\sum_{i=1}^{k}\lambda_i=1$, and $\lambda_i\geq 0$ for $i\in [k]$. The spanning-tree polytope is integral so there are polynomial-time algorithms to construct this decomposition (See Proposition~\ref{cara} and surrounding discussion).

Let $O_i$ be the set of odd-degree vertices of spanning tree $T_i$.
%Consider the incidence vectors of feasible $O_i$-joins.
% \cindy{Arash, do you also want to replace this longer discussion with a reference?}
Adding an $O_i$-join to tree $T_i$ gives a solution to 2EC.
% CAP: TODO, use later?
%Let $S$ be any subset of vertices with an even number of odd-degree vertices.  The $O_i$-join LP requires at least one edge joining $S$ to the rest of the vertices. The subtour-elimination LP requires every subset of vertices have two edges connecting it to the rest of the vertices.  Therefore, since $x$ is a feasible solution to the subtour-elimination LP, $x/2$ is feasible for the $O_i$-join LP associated with spanning tree $T_i$.
\arash{I still think this following sentence is confusing: we are not using the observation of Wolsey. Actually I don't think Wolsey spefically talks about writing 3/2x as a convex combination of solutions... I think it's would be better if we just explain how $(\frac{|V|-1}{|V|} + \alpha) x$ as convex combination of solutions and cite Wolsey as a general thing. How about deleting the next sentence, and after describing the LP below saying that $y^i = \frac{x}{2}$ and $\alpha = \frac{1}{2}$ is a feasible solution to this LP yielding the integarlity gap of $\frac{3}{2}$ of the subtour elimination relaxation for the TSP \cite{wolsey}.}
We will use Wolsey's observation~\cite{wolsey} that $\frac{x}{2}$ is in the convex hull of incidence vectors of the $O_i$-join of $G$ (denoted $O_i\join(G)$).
%\cindy{There was another instance of ``It's easy to see that...'' Given that (quite reasonably) you do not expect the reader to know what an O-join is, you should probably just give the short justification for this statement and remove ``It's easy to see that.''}
%\cindy{Do you have to precompute the joins (If they are more than the spanning trees themselves)?  Does that mean generating spanning trees before setting up the LP?  Or generating them on the fly?  Only minimum-weight spanning trees?  It seems that with a little more explanation, a reader who has not been immersed in 2EC and TSP could follow this.}
We solve the following LP that allows us to find edges that simultaneously augment all the spanning trees to solutions to 2EC.
%$O$-join-based parity corrections that are good for the whole convex combination.
\begin{equation}\label{ojoinaverage}
\min \{ \alpha:\;\sum_{i=1}^{k} \lambda_i y^i = \alpha \cdot x,\;  y^i \in \dom(O_i\join(G)) \; \mbox{for $i\in [k]$}\}.
\end{equation}
The variables in the above LP are $y^i\in \mathbb{R}^{E}_{\geq 0}$ for $i\in [k]$, where $E$ is the number of edges in the graph.  The parameters $\lambda_i$ are those we computed in~(\ref{eq:tree-decomp}). For each $i\in [k]$, we require $y^i$ to be in the dominant of $O_i\join(G)$.  We enforce this by adding a set of constraints for each $i$ from~\cite[p. 490]{schrijverbook}. These constraints force the inclusion of at least one edge from the cut associated with each set of vertices that have an even number of elements from $O_i$. The $O_i$ joins are selected fractionally \arash{(specifically $y^i$ need not be integer vectors)}, but because the $O_i$-join polytope is integral, we can \arash{replace $y^i$ with a convex combination of} find $O_i$-joins of $G$ (proposition~\ref{cara}).  This method gives a ($\frac{|V|-1}{|V|}+\alpha$)-approximation for the specific instance $x$. In general this can be less than the classic $\frac{3}{2}$.
%\cindy{You say the $y^i$ are LP variables, not integers. It even looks like they can even be more than $1$. What does it mean for a non-binary vector to be in a set of binary variables?  I would have expected the lambdas to be the variables, given their use as the convex combination multipliers in most of the paper.}
%This formulation allows the instance-specific approximation ratio of Christofides' algorithm to be below $\frac{3}{2}$.
\cindy{Arash, I liked your explanation about how a single edge will need to contribute 3/2 in the classic method.  You can add that if you'd like, but that might not be necessary given the time.}
